---
title: "covid 19"
author: "Martha Leal Olmedo"
date: "11/12/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
In the two bar charts above, the first bar chart shows there were more females than males in the alive tab and there was more males in the dead tab compared to the females. The second bar chart shows a great increase in death than the first bar chart. And in the second bar chart, the males death is higher than the females death. So the first plot shows that the amount of people that died was not that great compared to the people who are still alive, the second plot shows that there is more dead males than the dead females.



  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

In the tables presented, I see that both males and females have the total of 100. I see that on the top table the males alive is 44 and the males dead are 64. It also shows that the females alive are 56 and the females dead are 36. The other table shows that 98.7% females are alive, 1.4 are dead. And for the males it shows 97% are alive and 3 are dead. The total for the females in this table is 100.0.



## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The Chi sq test show that 1831 were females and that 1411 were males that were alive, the chi sq also shows that 25 females and 44 males died. The fisher test show me information of the odds, for the females dying which is 25/1831 and for males it is 44/1411. The odd ratio is 44/1411 divided by 25/1831 and the answer is 2.3.The odd ratio shows me that the males are 2.3x more likely to die than the females are too.


## Conclusions

In conclusion the data shows that males are more likely to die from covid-19 compared to the females. Since the data was taken at the start of covid, there is very little information and not a lot was available to us.
