---
title: "covid 19"
author: "Martha Leal Olmedo"
date: "11/12/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
In the two bar charts above, the first bar chart shows there were more females than males in the alive tab and there was more males in the dead tab compared to the females. The second bar chart shows a great increase in death than the first bar chart. And in the second bar chart, the males death is higher than the females death. So the first plot shows that the amount of people that died was not that great compared to the people who are still alive, the second plot shows that there is more dead males than the dead females.



  
## Numerical results


## Inferential results 




## Conclusions

